---
title: "Download data from the database"
output: rmarkdown::github_document
author: Tristan Mahr
date: "`r Sys.Date()`"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  fig.path = "figs/00-",
  echo = TRUE, 
  comment = "#>", 
  collapse = TRUE,
  fig.width = 6, 
  fig.asp = 0.62)
```

This script downloads response-level data from the minimal pairs
experiment from timepoint 1 of the longitudinal study.

The data-set includes only trials from the most frequently used
item-pairs. (There was some variation in the item counts and item sets
in the initial versions of the task, so we elect to keep the most
frequent items.) We exclude a participant with too few items from their
administration of the experiment. (Perhaps a crashed or aborted
experiment.) We exclude a participant who apparently chose the wrong
item on every trial. We also include EVT-2, PPVT-4 and GFTA-2 as these
are useful covariates.


Connect and download the scores
------------------------------------------------------------------------

Connect to UMN database using the L2TDatabase package.
<https://github.com/LearningToTalk/L2TDatabase>

```{r}
# Download test scores from the database
library(L2TDatabase)
library(dplyr)
library(ggplot2)

# Connect to database. This assumes the L2T Database configuration file 
# is the user's home folder.
db_config_file <- path.expand("~/l2t_db.cnf")
l2t <- l2t_connect(db_config_file, "backend")

# Get the timepoint 1 minimal pair responses and vocabulary scores. (I'm
# using the backend database because I wrote the first version of this
# script using it.)
min_pair_raw <- tbl(l2t, "MinPair_Admin") %>%
  left_join(tbl(l2t, "MinPair_Responses")) %>%
  left_join(tbl(l2t, "ChildStudy")) %>%
  left_join(tbl(l2t, "Child")) %>%
  left_join(tbl(l2t, "Study")) %>%
  left_join(tbl(l2t, "EVT")) %>% 
  left_join(tbl(l2t, "PPVT")) %>% 
  left_join(tbl(l2t, "GFTA")) %>% 
  # No practice trials
  filter(Study == "TimePoint1", Running == "Test") %>%
  select(
    Study, ResearchID = ShortResearchID, Female:CImplant,
    MinPair_Age, MinPair_Dialect:Correct,
    PPVT_Age, PPVT_Raw:PPVT_GSV,
    EVT_Age, EVT_Raw:EVT_GSV,
    GFTA_Age, GFTA_RawCorrect:GFTA_Standard) %>%
  select(
    -MinPair_Completion, -MinPair_Notes, 
    -MinPair_Admin_Timestamp, -ResponseID) %>% 
  collect(n = Inf)

min_pair_raw
```


Screen the data
------------------------------------------------------------------------

```{r}
# Function to sort two words, combine with hyphen.
# c("dog", "cat") -> "cat-dog"
make_word_pair <- function(x, y) {
  paste0(sort(c(x, y)), collapse = "-")
}

# Use the word-pair function on each row
min_pair_raw <- min_pair_raw %>%
  mutate(WordPair = purrr::map2_chr(Item1, Item2, make_word_pair))

# Keep only the word-pairs used in every administration. With ~190
# children, these are words with at least 300 responses.
frequent_items <- min_pair_raw %>%
  count(WordPair) %>%
  filter(300 < n) %>%
  print() %>%
  select(WordPair)

# Who is missing the common items?
too_few_trials <- min_pair_raw %>%
  inner_join(frequent_items) %>%
  count(ResearchID) %>%
  filter(n != 30) %>%
  print()

# Who is that child who responded with the opposite answer?
impossibly_low <- min_pair_raw %>%
  group_by(ResearchID) %>%
  summarise(PropCorrect = mean(Correct)) %>%
  filter(PropCorrect < .15)
pbinom(2, 30, .5)

freq_items <- min_pair_raw %>%
  inner_join(frequent_items) %>%
  anti_join(too_few_trials) %>%
  filter(CImplant == 0)

prop_correct <- freq_items %>%
  group_by(ResearchID) %>%
  summarise(MeanCorrect = mean(Correct), n = n())
```


Make some plots
------------------------------------------------------------------------

```{r minp-raw-histogram-with-outlier}
p1 <- ggplot(prop_correct) +
  aes(x = MeanCorrect) +
  # bins are 5% quantiles
  geom_histogram(binwidth = .05, center = .525, color = "white") +
  xlab("Observed proportion of items correct") +
  ylab("N participants") +
  theme_grey(base_size = 16) +
  annotate(
    "label", x = .15, y = 5, 
    label = "Chose opposite\nitem on purpose", size = 4.5)
p1
```

```{r min-raw-histogram}
p2 <- ggplot(prop_correct %>% anti_join(impossibly_low)) +
  aes(x = MeanCorrect) +
  # bins are 5% quantiles
  geom_histogram(binwidth = .05, center = .525, color = "white") +
  xlab("Observed proportion of items correct") +
  ylab("N participants") +
  theme_grey(base_size = 16)
p2
```


Save data (without outlier)
------------------------------------------------------------------------

```{r}
freq_items %>%
  anti_join(impossibly_low) %>%
  readr::write_csv("./data/raw-minimal-pairs-responses.csv")
```
